---
title: "Homework 6"
author: "Jenesis Merriman"
date: "December 3, 2022"
output: github_document
---

```{r setup, include=FALSE} 
library(tidyverse) #used
library(broom) #used
library(readr) #used
library(p8105.datasets) #delete unnecessary packages
library(mgcv)
library(modelr)

library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1


## Problem 2

This problem uses data gathered by the Washington Post on homicides in 50 large U.S. cities.

First, I will load and tidy the data. The following code uses `mutate` to update variables to appropriate types, convert reported_date into a more readable format (YYY-MM-DD), and correct the capitalization of victim_first and victim_last. It also creates two new variables: i) a `city_state` variable that returns both city and state (e.g. “Baltimore, MD”), and ii) a binary variable `solved` indicating whether the homicide is solved (TRUE) or not (FALSE). Solved homicides are defined as homicides for which the disposition is "Closed by arrest." `filter` is used to omit homicides in Dallas, TX; Phoenix, AZ; Kansas City, MO; and Tulsa, AL, and to limit our analysis to homicides for which victim_race is white or Black.

```{r}
homicides_hw6 =
  read_csv("./data/homicide-data.csv") %>% #load data
  janitor::clean_names() %>% #clean names
  mutate(city_state = as.factor(str_c(city, state, sep = ", ")), #new variable
         solved = as.logical(ifelse(disposition %in% c("Closed by arrest"), TRUE, FALSE)), #new variable
         reported_date = as.Date(as.character(reported_date),"%Y%m%d"), #fixes date format
         victim_age = as.numeric(victim_age), #character to double
         victim_first = str_to_title(victim_first), #fixes all caps
         victim_last = str_to_title(victim_last), #fixes all caps
         victim_sex = as.factor(victim_sex), #character to factor
         victim_race = as.factor(victim_race), #character to factor
         city = as.factor(city), #character to factor
         state = as.factor(state)) %>% #character to factor
  filter(city_state != "Dallas, TX" & city_state != "Phoenix, AZ" & city_state != "Kansas City, MO" & city_state != "Tulsa, AL") %>%
  filter(victim_race == "White" | victim_race == "Black")
```

Next, I will use the `glm` function to fit a logistic regression with solved vs unsolved as the outcome and victim age, sex and race as predictors for the city of Baltimore, MD.

```{r}
baltimore =
  homicides_hw6 %>%
  filter(city_state == "Baltimore, MD")

fit = glm(solved ~ victim_age + victim_sex + victim_race, data = baltimore, family = "binomial")
```

The following code saves the output of my logistic regression model as an R object. `broom::tidy` and `filter` are used to obtain the adjusted odds ratio estimate and confidence interval for solving homicides comparing male victims to female victims, keeping all other variables fixed.

```{r}
fit_output =
  fit %>% 
  tidy(exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95) %>%
  filter(term == "victim_sexMale") %>%
  select(term, "adjusted_OR" = "estimate", "CI_lower" = "conf.low", "CI_upper" = "conf.high")

fit_output
```

Now, I will run `glm` for each of the cities in the homicides_hw6 dataset, and extract the adjusted odds ratio and confidence interval for solving homicides comparing male victims to female victims. The following "tidy" pipeline uses `purrr::map`, `glm`, `broom::tidy`, list columns, and `unnest` to create a dataframe with estimated ORs and CIs for each city. `filter` and `select` are used to select our values of interest.

```{r}
city_glm =
  homicides_hw6 %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(data, ~glm(solved ~ victim_age + victim_sex + victim_race, family = "binomial", data = .x)),
    exp_results = map(models, tidy, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95)) %>%
  select(-data, -models) %>% 
  unnest(exp_results) %>%
  filter(term == "victim_sexMale") %>%
  select(city_state, 
         term, 
         "adjusted_OR" = "estimate", 
         "CI_lower" = "conf.low",
         "CI_upper" = "conf.high")
city_glm
```

Finally, the following code uses `geom_point` and `geom_errorbar` to create a plot showing the estimated odds ratios and confidence intervals for each city. Cities are arranged in ascending order of estimated OR.

```{r}
city_glm %>%
  ggplot(aes(x = reorder(city_state, +adjusted_OR), y = adjusted_OR)) +
  geom_point(show.legend = FALSE) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
    title = "Adjusted OR for homicide resolution comparing male victims to female victims",
    x = "City",
    y = "Adjusted OR")
```
 
 **Comment:** The plot suggests that the odds of homicide resolution for male victims is less than the odds of homicide resolution for female victims in most of our analyzed cities, in general, after adjusting for victim age and race.In other words, after adjusting for victim age and race, homicides are generally more likely to be solved in cases where the victim is a woman compared to cases where the victim is a man. This is true for all cities in our dataset **except** Atlanta, GA; Richmond, VA; Nashville, TN; Fresno, CA; Stockton, CA; and Albuquerque, NM, where the adjusted OR >= 1, which indicates that the odds of homicide resolution for male victims is equal to or greater than the odds of homicide resolution for female victims in these cities.

## Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children.

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
birthweight =
  read_csv("./data/birthweight.csv") %>% #load data
  janitor::clean_names() %>%
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace))

colSums(is.na(birthweight)) #check for missing data: none missing
lapply(birthweight, unique) #check for unusual values: menarche = 0 (1 row)
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process.

```{r modeling}
fit_step = step(lm(bwt ~ ., birthweight), direction = "both", trace = FALSE)
tidy(fit_step)
```

Show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r residuals plot}
birthweight %>%
  modelr::add_residuals(fit_step) %>%
  modelr::add_predictions(fit_step) %>%
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  geom_hline(yintercept = 0,linetype = "dashed")
```


Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)
* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
fit_i = lm(bwt ~ blength + gaweeks, data = birthweight)
fit_ii = lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex, data = birthweight)

tidy(fit_i)
tidy(fit_ii)
```

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.





Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

